### 2020-09-19 manuscript/ERL_global_C_review.Rmd
worked on writing, refs

*Review Results/ Synthesis.*  ForC v.XX yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets. The rate of C cycling generally increased from boreal to tropical regions, whereas C stocks showed less directional variation. The majority of flux variables, together with most live biomass pools, increased significantly with stand age, and the rate of increase again tended to increase from boreal to tropical regions.

*Discussion.* This analysis yields a comprehensive, largely consistent picture of C cycling across the world's forests.   *[Discussion section will interpret results, highlighting new and significant findings, and discuss implications. Tentative headings are “Stand level C cycling in forests globally”, “Age trends in C cycling”, and “Implications for climate change mitigation”.]*

@@

Forest ecosystems globally play a critical role in regulating atmospheric carbon dioxide (CO~2~), and thereby will shape the course of climate change (IPCC1.5). 
Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**). 
While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades. 
The resulting CO~2~ sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). 
Moreover, forests contain substantial carbon (C) stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C [biomass and soils; @bonan_forests_2008]. 
Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**). 
The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. 
Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017; Cavaleri et al 2015, Griscom et al 2017].

Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [Friedlingstein et al 2006; @krause_large_2018; REFS] and CO~2~ accounting frameworks [Pan, REFS] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history. 
These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published. 
While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [@li_mapping_2019; **REFS for biomass, biomass change, net CO2 flux**], measurement of most forest C stocks and fluxes require intensive on-the-ground data collection. 
Original studies typically cover only a small numbers of sites at a time, with rare exceptions spanning regions or continents [e.g., @lutz_global_2018; FLUXNET_REF), typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015] or FLUXNET [@baldocchi_fluxnet_2001].
The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*. 
In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle (*e.g.*, Pan et al 2011), or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere (REFS).

To address the need for global-scale analyses of forest C cycling, we recently developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018). 
ForC contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables) based on ground-based measurements, along with associated data required for interpretation (*e.g.*, stand history, measurement methods).
These data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations, with notable recent additions of the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020]. 
ForC currently contains **#** (~49,000!) records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

Here, we synthesize ForC data (Fig. 1) to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address three broad questions:

@@

This review synthesizes data from the ForC database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018]. 
ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies. 
Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records. 
The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. 
As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018]. 
Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018]. 
Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database (GROA v1.0; Cook-Patton et al. in press; ZenodoDOI), and data from several publications (Taylor et al. 2017, Lutz et al. 2018, Johnson_2018_csss; see GitHub repo for complete list) ([GitHub list](https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv)). 
We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: TBD).

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature). 
Climate zones were defined according to Köppen-Geiger zones, which were extracted based on site geographic coordinates from the ESRI Köppen-Geiger map (downloaded June 2014 from http://maps3.arcgisonline.com/ArcGIS/rest/services/A-16/Köppen- Geiger_Observed_and_Predicted_Climate_Shifts/MapServer; Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd). 
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations (SYNMAP; Jung et al 2006) and recorded in the  dominant.veg field of the ForC MEAUREMENTS table. 
Stand age was obtained from the stand.age field of MEASUREMENTS table, and forests classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
Records for which stand age was unknown were excluded from the analysis. 
These groupings were defined based on ForC v.2.0 [@anderson-teixeira_forc_2018], for which they covered 91% of the primary variable records for forests of known age, or 86% of total records. 
The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.

>>>>

*Review Results/ Synthesis.*  ForC v.XX yielded a fairly comprehensive picture of C cycling in the world's major forest biomes, with broad closure of C budgets. The rate of C cycling generally increased from boreal to tropical regions, whereas C stocks showed less directional variation. The majority of flux variables, together with most live biomass pools, increased significantly with stand age, *and the rate of increase again tended to increase from boreal to tropical regions*.

*Discussion.* This analysis yields a comprehensive and broadly consistent picture of C cycling across the world's forests.   *[Discussion section will interpret results, highlighting new and significant findings, and discuss implications. Tentative headings are “Stand level C cycling in forests globally”, “Age trends in C cycling”, and “Implications for climate change mitigation”.]*

@@

Forest ecosystems globally play a critical role in regulating atmospheric carbon dioxide (CO~2~), and thereby will shape the course of climate change (IPCC1.5). 
Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**). 
While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades. 
The resulting CO~2~ sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). 
Moreover, forests contain substantial carbon (C) stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C [biomass and soils; @bonan_forests_2008]. 
Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), reducing the net forest sink to ~1.2-1.7 Gt C yr^-1^ across Earth's forests [Le Quéré et al 2017, @schimel_observing_2015] (**update, give range**). 
The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. 
Regrowth forests in particular will play an important role [@pugh_role_2019], as these represent a large (~#%) and growing proportion of Earth's forests.
Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; Cavaleri et al 2015].

Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climate-carbon_2006; @krause_large_2018; REFS] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history. 
These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published. 
While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [@li_mapping_2019; **REFS for biomass, biomass change, net CO2 flux**], measurement of most forest C stocks and fluxes require intensive on-the-ground data collection. 
Original studies typically cover only a small numbers of sites at a time, with rare exceptions spanning regions or continents [e.g., @lutz_global_2018; FLUXNET_REF), typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015] or FLUXNET [@baldocchi_fluxnet_2001].
The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*. 
In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere (REFS).

To address the need for global-scale analyses of forest C cycling, we recently developed an open-access Global Forest Carbon database, ForC (@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018). 
ForC contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables) based on ground-based measurements, along with associated data required for interpretation (*e.g.*, stand history, measurement methods).
These data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
Since the its most recent publication [@anderson-teixeira_forc_2018], ForC has been integrated with two large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020], both of which have also synthesized published forest C data. 
Following these additions, ForC currently contains **#** (~49,000!) records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

Here, we synthesize ForC data (Fig. 1) to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. Our primary goal is to provide a data-based summary of our current state of knowledge on broad trends in forest C cycling. We address three broad questions:

@@

This review synthesizes data from the ForC database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018]. 
ForC amalgamates numerous intermediary data sets (*e.g.*, REFS) and original studies. 
Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records. 
The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age. 
As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018]. 
Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], yet these represented only approximately one-third of records in ForC v.2.0 [@anderson-teixeira_forc_2018]. 
Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [SRDB v.##; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [GROA v1.0; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020], and data from several publications [e.g., @taylor_temperature_2017, @lutz_global_2018, @johnson_climate_2018; detailed list at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv]. 
We note that there remains a significant amount of relevant data that is not yet included in ForC, particularly biomass data from national forest inventories (*e.g.*,: REFS). 
The database version used for this analysis has been tagged as a new release on Github (XX) and assigned a DOI through Zenodo (DOI: TBD).

For this analysis, we grouped forests into four broad biome types (tropical broadleaf, temperate broadleaf, temperate needleleaf, and boreal needleleaf) and two age classifications (young and mature). 
Climate zones (Fig. 1) were defined based on site geographic coordinates according to Köppen-Geiger zones (Rubel and Kottek 2010).
Tropical climates were defined to include all equatorial (A) zones, temperate climates were defined to include all warm temperate (C) zones and warmer snow climates (Dsa, Dsb, Dwa, Dwb, Dfa, and Dfb), and boreal climates were defined to include the colder snow climates (Dsc, Dsd, Dwc, Dwd, Dfc, and Dfd). 
Any forests in dry (B) and polar (E) Köppen-Geiger zones were excluded from the analysis. 
Leaf type (broadleaf / needleleaf) was defined based on descriptions in original publications (prioritized) or values extracted from a global map based on satellite observations [SYNMAP; @jung_exploiting_2006]. 
Forests were classified as “young” (< 100 years) or “mature” ($\ge$ 100 years or classified as “mature”, “old growth”, “intact”, or “undisturbed” in original publication). 
Records for which stand age was unknown (#% of total database) were excluded from the analysis. 
*These groupings were defined based on ForC v.2.0 [@anderson-teixeira_forc_2018], for which they covered 91% of the primary variable records for forests of known age, or 86% of total records. The most well-represented forest types excluded were boreal broadleaf and boreal and temperate mixed broadleaf-needleleaf, each with <400 records total for stands of any age.*
